---
title: "3_modeling"
author: "Joy_Fu"
date: "2023-05-02"
output: html_document
---

```{r message=FALSE, warning=FALSE}
rm(list = ls())
# lapply(paste('package:', names(sessionInfo()$otherPkgs), sep = ""),
#        detach, character.only = TRUE, unload = TRUE)
pacman::p_load(tidyverse, h2o, epiflow, caret)
raw_data_path = "/Users/Mingzhou/Desktop/Projects/Dementia-prediction/data/"
output_path = "/Users/Mingzhou/Desktop/Projects/Dementia-prediction/output/"
# Source in useful functions
source("/Users/Mingzhou/Desktop/Projects/Dementia-prediction/code/functions.R")
# Load in PRS + geno beta data
load(file = paste0(raw_data_path, "prs/mod/atlas_prs_final.rda"))
load(file = paste0(raw_data_path, "modeling/beta_merge_raw.rda"))
load(file = paste0(raw_data_path, "modeling/freeze60k_geno_freq_full_SNP.rda"))
# Read in SNP.id data
ids_folder = paste0(raw_data_path, "modeling/snp_ids") 
ids_files = list.files(ids_folder, pattern = ".rda$")
for (i in ids_files) {
  load(paste0(raw_data_path, 'modeling/snp_ids/', i))
}
```

# Part 1. AMR analyses
```{r}
# Read in cleaned data
load(file = paste0(raw_data_path, "modeling/AMR/sample_AMR_final.rda"))
# Join PRSs
sample_AMR_prs = sample_AMR_final %>% 
  mutate(UniqueSampleId = as.character(UniqueSampleId)) %>% 
  left_join(atlas_prs_final) %>% column_to_rownames(var = "UniqueSampleId") %>% 
  select(-c(gen_ancestry, APOE, e2count)) %>% drop_na()
dim(sample_AMR_prs) # dim = (1638,29)
# Set response variable
Y = "dementia"
demo_set = c("age", "female", paste0('PC', 1:4))
# set h2o project
# h2o.init(nthreads = -1)
```

## 1. PRS models
### 1) Model setups
```{r}
target.train = as.h2o(sample_AMR_prs)
target.train[,Y] = as.factor(target.train[,Y])
```

### 2) Single PRS models
```{r}
seed_num = 20224766
#========= Benchmark: demo only =========
X = c("age", "female")
demo.only = h2o.glm(training_frame = target.train, x = X, y = Y, 
                    balance_classes = T, model_id = "Demo only", 
                    nfolds = 5, fold_assignment = "Stratified",
                    seed = seed_num, family = "binomial", lambda = 0, 
                    standardize = T, keep_cross_validation_models = F)
#========= single PRS model =========
# AD PRS EUR
X = c("AD_PRS_EUR_indsig", demo_set)
PRS.EUR.indsig = h2o.glm(training_frame = target.train, x = X, y = Y, 
                         balance_classes = T, model_id = "AD PRS EUR indsig", 
                         nfolds = 5, fold_assignment = "Stratified", 
                         seed = seed_num, family = "binomial", lambda = 0, 
                         standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_lead", demo_set)
PRS.EUR.lead = h2o.glm(training_frame = target.train, x = X, y = Y, 
                       balance_classes = T, model_id = "AD PRS EUR lead", 
                       nfolds = 5, fold_assignment = "Stratified", 
                       seed = seed_num, family = "binomial", lambda = 0, 
                       standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_map", demo_set)
PRS.EUR.map = h2o.glm(training_frame = target.train, x = X, y = Y, 
                      balance_classes = T, model_id = "AD PRS EUR map", 
                      nfolds = 5, fold_assignment = "Stratified", 
                      seed = seed_num, family = "binomial", lambda = 0, 
                      standardize = T, keep_cross_validation_models = F)
# AD PRS AFR
X = c("AD_PRS_AFR_indsig", demo_set)
PRS.AFR.indsig = h2o.glm(training_frame = target.train, x = X, y = Y, 
                         balance_classes = T, model_id = "AD PRS AFR indsig", 
                         nfolds = 5, fold_assignment = "Stratified", 
                         seed = seed_num, family = "binomial", lambda = 0, 
                         standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_AFR_lead", demo_set)
PRS.AFR.lead = h2o.glm(training_frame = target.train, x = X, y = Y, 
                       balance_classes = T, model_id = "AD PRS AFR lead", 
                       nfolds = 5, fold_assignment = "Stratified", 
                       seed = seed_num, family = "binomial", lambda = 0, 
                       standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_AFR_map", demo_set)
PRS.AFR.map = h2o.glm(training_frame = target.train, x = X, y = Y, 
                      balance_classes = T, model_id = "AD PRS AFR map", 
                      nfolds = 5, fold_assignment = "Stratified", 
                      seed = seed_num, family = "binomial", lambda = 0, 
                      standardize = T, keep_cross_validation_models = F)
# AD PRS Trans 
X = c("AD_PRS_Trans_indsig", demo_set)
PRS.trans.indsig = h2o.glm(training_frame = target.train, x = X, y = Y, 
                           balance_classes = T, model_id = "AD PRS Trans indsig", 
                           nfolds = 5, fold_assignment = "Stratified", 
                           seed = seed_num, family = "binomial", lambda = 0, 
                           standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_Trans_lead", demo_set)
PRS.trans.lead = h2o.glm(training_frame = target.train, x = X, y = Y, 
                         balance_classes = T, model_id = "AD PRS Trans lead", 
                         nfolds = 5, fold_assignment = "Stratified", 
                         seed = seed_num, family = "binomial", lambda = 0, 
                         standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_Trans_map", demo_set)
PRS.trans.map = h2o.glm(training_frame = target.train, x = X, y = Y, 
                        balance_classes = T, model_id = "AD PRS Trans map", 
                        nfolds = 5, fold_assignment = "Stratified", 
                        seed = seed_num, family = "binomial", lambda = 0, 
                        standardize = T, keep_cross_validation_models = F)
# e4count
X = c("e4count", demo_set)
e4count = h2o.glm(training_frame = target.train, x = X, y = Y, 
                  balance_classes = T, model_id = "e4count", 
                  nfolds = 5, fold_assignment = "Stratified", 
                  seed = seed_num, family = "binomial", lambda = 0, 
                  standardize = T, keep_cross_validation_models = F)
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("demo.only", "e4count", 
               "PRS.EUR.indsig", "PRS.EUR.lead", "PRS.EUR.map", 
               "PRS.AFR.indsig", "PRS.AFR.lead", "PRS.AFR.map", 
               "PRS.trans.indsig", "PRS.trans.lead", "PRS.trans.map")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>%
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
single_prs_model_metric = rbind(all_model_metric)
single_prs_model_performance = rbind(all_model_performance)
single_prs_model_summary = single_prs_model_metric %>% 
  inner_join(single_prs_model_performance)
```

### 3) Multiple PRS model
```{r}
#========= multiple PRS model =========
# AD PRS combinations
X = c("AD_PRS_EUR_indsig", "AD_PRS_Trans_indsig", "AD_PRS_AFR_indsig", demo_set)
PRS.AD.indsig.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                balance_classes = T, 
                                model_id = "AD PRS ind.sig combine", 
                                nfolds = 5, fold_assignment = "Stratified", 
                                seed = seed_num, family = "binomial", lambda = 0, 
                                standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_lead", "AD_PRS_Trans_lead", "AD_PRS_AFR_lead", demo_set)
PRS.AD.lead.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                              balance_classes = T, 
                              model_id = "AD PRS lead combine", 
                              nfolds = 5, fold_assignment = "Stratified", 
                              seed = seed_num, family = "binomial", lambda = 0, 
                              standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_map", "AD_PRS_Trans_map", "AD_PRS_AFR_map", demo_set)
PRS.AD.map.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                             balance_classes = T, 
                             model_id = "AD PRS map combine", 
                             nfolds = 5, fold_assignment = "Stratified", 
                             seed = seed_num, family = "binomial", lambda = 0, 
                             standardize = T, keep_cross_validation_models = F)
# AD + neuro combinations
X = c("AD_PRS_EUR_indsig", "AD_PRS_Trans_indsig", "AD_PRS_AFR_indsig", 
      "PD_PRS_indsig", "LBD_PRS_indsig", "PSP_PRS_indsig", 
      "Stroke_PRS_indsig", demo_set)
PRS.neuro.indsig.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                   balance_classes = T, 
                                   model_id = "Neuro PRS ind.sig combine", 
                                   nfolds = 5, fold_assignment = "Stratified", 
                                   seed = seed_num, family = "binomial", lambda = 0, 
                                   standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_lead", "AD_PRS_Trans_lead", "AD_PRS_AFR_lead", 
      "PD_PRS_lead", "LBD_PRS_lead", "PSP_PRS_lead", 
      "Stroke_PRS_lead", demo_set)
PRS.neuro.lead.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                 balance_classes = T, 
                                 model_id = "Neuro PRS lead combine", 
                                 nfolds = 5, fold_assignment = "Stratified", 
                                 seed = seed_num, family = "binomial", lambda = 0, 
                                 standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_map", "AD_PRS_Trans_map", "AD_PRS_AFR_map", 
      "PD_PRS_map", "LBD_PRS_map", "PSP_PRS_map", "Stroke_PRS_map", demo_set)
PRS.neuro.map.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                balance_classes = T,
                                model_id = "Neuro PRS map combine", 
                                nfolds = 5, fold_assignment = "Stratified", 
                                seed = seed_num, family = "binomial", lambda = 0, 
                                standardize = T, keep_cross_validation_models = F)
# LASSO selected PRSs
X = c(demo_set, names(target.train)[8:28])
PRS.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = 1, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
prs_select = PRS.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
# Step 2: fit LASSO
X = c(prs_select, demo_set)
PRS.neuro.select = h2o.glm(training_frame = target.train, x = X, y = Y, 
                           balance_classes = T, model_id = "LASSO PRSs", 
                           nfolds = 5, fold_assignment = "Stratified", 
                           seed = seed_num, family = "binomial",
                           alpha = 0, lambda_search = TRUE,
                           standardize = T, keep_cross_validation_models = F)
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("PRS.AD.indsig.combine", "PRS.AD.lead.combine", 
               "PRS.AD.map.combine", "PRS.neuro.indsig.combine", 
               "PRS.neuro.lead.combine", "PRS.neuro.map.combine", 
               "PRS.neuro.select")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
multi_prs_model_metric = rbind(all_model_metric)
multi_prs_model_performance = rbind(all_model_performance)
multi_prs_model_summary = multi_prs_model_metric %>% 
  inner_join(multi_prs_model_performance)

full_prs_model_summary = rbind(single_prs_model_summary, 
                               multi_prs_model_summary) %>% as.data.frame()
```

## 2. Weighted-SNP models
```{r}
snp_df_full = sample_AMR_final %>% 
  mutate(UniqueSampleId = as.character(UniqueSampleId)) %>% 
  left_join(genotype_df_fam, by = c("UniqueSampleId" = "IID")) %>% 
  select(-c(UniqueSampleId)) 
# LASSO selected SNPs
target.train = as.h2o(snp_df_full)
target.train[,Y] = as.factor(target.train[,Y])
# Set parameters
alpha_feature = 0.6
```

### Run two-step linear models
```{r}
# AD indsig SNPs
# Step 1: feature selection
X = c(demo_set, AD.indsig.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AMR_AD_indsig = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AMR_AD_indsig))
# Step 2: fit LASSO
X = c(snps_AMR_AD_indsig, demo_set)
aml_AD_indsig = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.AD.indsig = aml_AD_indsig@leader

# AD lead SNPs
# Step 1: feature selection
X = c(demo_set, AD.lead.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AMR_AD_lead = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AMR_AD_lead))
# Step 2: fit LASSO
X = c(snps_AMR_AD_lead, demo_set)
aml_AD_lead = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.AD.lead = aml_AD_lead@leader

# AD map SNPs
# Step 1: feature selection
X = c(demo_set, AD.map.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AMR_AD_map = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AMR_AD_map))
# Step 2: fit LASSO
X = c(snps_AMR_AD_map, demo_set)
aml_AD_map = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.AD.map = aml_AD_map@leader

# All indsig SNPs
# Step 1: feature selection
X = c(demo_set, all.indsig.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AMR_all_indsig = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AMR_all_indsig))
# Step 2: fit LASSO
X = c(snps_AMR_all_indsig, demo_set)
aml_all_indsig = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.all.indsig = aml_all_indsig@leader

# All lead SNPs
# Step 1: feature selection
X = c(demo_set, all.lead.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AMR_all_lead = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AMR_all_lead))
# Step 2: fit LASSO
X = c(snps_AMR_all_lead, demo_set)
aml_all_lead = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.all.lead = aml_all_lead@leader

# All mapped SNPs
# Step 1: feature selection
X = c(demo_set, all.map.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AMR_all_map = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AMR_all_map))
# Step 2: fit LASSO
X = c(snps_AMR_all_map, demo_set)
aml_all_map = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.all.map = aml_all_map@leader
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("snp.lasso.AD.indsig", "snp.lasso.AD.lead", 
               "snp.lasso.AD.map", "snp.lasso.all.indsig",
               "snp.lasso.all.lead", "snp.lasso.all.map")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
lasso_snp_model_metric = rbind(all_model_metric)
lasso_snp_model_performance = rbind(all_model_performance)
lasso_snp_model_summary = lasso_snp_model_metric %>% 
  inner_join(lasso_snp_model_performance)
```

## 3. AutoML
```{r warning=FALSE}
# First use the sample selection results from LASSO 
X = c(snps_AMR_all_map, demo_set)
# Run AutoML for 10 base models (limited to 1 hour max runtime by default)
aml.full.map = h2o.automl(x = X, y = Y, training_frame = target.train, 
                          balance_classes = TRUE, max_models = 10, 
                          seed = seed_num, sort_metric = "AUCPR")
# Optionally edd extra model information to the leaderboard
lb.aml = h2o.get_leaderboard(aml.full.map, extra_columns = "ALL")
# Print all rows (instead of default 6 rows)
print(lb.aml, n = nrow(lb.aml))
```

```{r}
# Get the "best" for each type of model
best_gbm = h2o.getModel("GBM_1_AutoML_228_20230615_230614")
best_xgboost = h2o.getModel("XGBoost_3_AutoML_228_20230615_230614")
best_drf = h2o.getModel("DRF_1_AutoML_228_20230615_230614")
best_xrt = h2o.getModel("XRT_1_AutoML_228_20230615_230614")
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("best_gbm", "best_xgboost", "best_drf", "best_xrt")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
aml_snp_model_metric = rbind(all_model_metric)
aml_snp_model_performance = rbind(all_model_performance)
aml_snp_model_summary = aml_snp_model_metric %>% 
  inner_join(aml_snp_model_performance)
```

```{r}
full_model_summary = rbind(full_prs_model_summary, 
                           lasso_snp_model_summary,
                           aml_snp_model_summary) %>% as.data.frame()
# Output results to table
write.table(full_model_summary, 
            file = paste0(output_path, "Model_performance_logistic_AMR.txt"),
            sep = "\t", quote = F, row.names = F, col.names = T)
# h2o.shutdown(prompt = F)
```

# Part 2. AFR analyses
```{r}
# Read in cleaned data
load(file = paste0(raw_data_path, "modeling/AFR/sample_AFR_final.rda"))
# Join PRSs
sample_AFR_prs = sample_AFR_final %>% 
  mutate(UniqueSampleId = as.character(UniqueSampleId)) %>% 
  left_join(atlas_prs_final) %>% column_to_rownames(var = "UniqueSampleId") %>% 
  select(-c(gen_ancestry, APOE, e2count)) %>% drop_na()
dim(sample_AFR_prs) # dim = (1092,29)
# Set response variable
Y = "dementia"
demo_set = c("age", "female", paste0('PC', 1:4))
# set h2o project
# h2o.init(nthreads = -1)
```

## 1. PRS models
### 1) Model setups
```{r}
target.train = as.h2o(sample_AFR_prs)
target.train[,Y] = as.factor(target.train[,Y])
```

### 2) Single PRS models
```{r}
seed_num = 2394
#========= Benchmark: demo only =========
X = c("age", "female")
demo.only = h2o.glm(training_frame = target.train, x = X, y = Y, 
                    balance_classes = T, model_id = "Demo only", 
                    nfolds = 5, fold_assignment = "Stratified",
                    seed = seed_num, family = "binomial", lambda = 0, 
                    standardize = T, keep_cross_validation_models = F)
#========= single PRS model =========
# AD PRS EUR
X = c("AD_PRS_EUR_indsig", demo_set)
PRS.EUR.indsig = h2o.glm(training_frame = target.train, x = X, y = Y, 
                         balance_classes = T, model_id = "AD PRS EUR indsig", 
                         nfolds = 5, fold_assignment = "Stratified", 
                         seed = seed_num, family = "binomial", lambda = 0, 
                         standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_lead", demo_set)
PRS.EUR.lead = h2o.glm(training_frame = target.train, x = X, y = Y, 
                       balance_classes = T, model_id = "AD PRS EUR lead", 
                       nfolds = 5, fold_assignment = "Stratified", 
                       seed = seed_num, family = "binomial", lambda = 0, 
                       standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_map", demo_set)
PRS.EUR.map = h2o.glm(training_frame = target.train, x = X, y = Y, 
                      balance_classes = T, model_id = "AD PRS EUR map", 
                      nfolds = 5, fold_assignment = "Stratified", 
                      seed = seed_num, family = "binomial", lambda = 0, 
                      standardize = T, keep_cross_validation_models = F)
# AD PRS AFR
X = c("AD_PRS_AFR_indsig", demo_set)
PRS.AFR.indsig = h2o.glm(training_frame = target.train, x = X, y = Y, 
                         balance_classes = T, model_id = "AD PRS AFR indsig", 
                         nfolds = 5, fold_assignment = "Stratified", 
                         seed = seed_num, family = "binomial", lambda = 0, 
                         standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_AFR_lead", demo_set)
PRS.AFR.lead = h2o.glm(training_frame = target.train, x = X, y = Y, 
                       balance_classes = T, model_id = "AD PRS AFR lead", 
                       nfolds = 5, fold_assignment = "Stratified", 
                       seed = seed_num, family = "binomial", lambda = 0, 
                       standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_AFR_map", demo_set)
PRS.AFR.map = h2o.glm(training_frame = target.train, x = X, y = Y, 
                      balance_classes = T, model_id = "AD PRS AFR map", 
                      nfolds = 5, fold_assignment = "Stratified", 
                      seed = seed_num, family = "binomial", lambda = 0, 
                      standardize = T, keep_cross_validation_models = F)
# AD PRS Trans 
X = c("AD_PRS_Trans_indsig", demo_set)
PRS.trans.indsig = h2o.glm(training_frame = target.train, x = X, y = Y, 
                           balance_classes = T, model_id = "AD PRS Trans indsig", 
                           nfolds = 5, fold_assignment = "Stratified", 
                           seed = seed_num, family = "binomial", lambda = 0, 
                           standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_Trans_lead", demo_set)
PRS.trans.lead = h2o.glm(training_frame = target.train, x = X, y = Y, 
                         balance_classes = T, model_id = "AD PRS Trans lead", 
                         nfolds = 5, fold_assignment = "Stratified", 
                         seed = seed_num, family = "binomial", lambda = 0, 
                         standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_Trans_map", demo_set)
PRS.trans.map = h2o.glm(training_frame = target.train, x = X, y = Y, 
                        balance_classes = T, model_id = "AD PRS Trans map", 
                        nfolds = 5, fold_assignment = "Stratified", 
                        seed = seed_num, family = "binomial", lambda = 0, 
                        standardize = T, keep_cross_validation_models = F)
# e4count
X = c("e4count", demo_set)
e4count = h2o.glm(training_frame = target.train, x = X, y = Y, 
                  balance_classes = T, model_id = "e4count", 
                  nfolds = 5, fold_assignment = "Stratified", 
                  seed = seed_num, family = "binomial", lambda = 0, 
                  standardize = T, keep_cross_validation_models = F)
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("demo.only", "e4count", 
               "PRS.EUR.indsig", "PRS.EUR.lead", "PRS.EUR.map", 
               "PRS.AFR.indsig", "PRS.AFR.lead", "PRS.AFR.map", 
               "PRS.trans.indsig", "PRS.trans.lead", "PRS.trans.map")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>%
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
single_prs_model_metric = rbind(all_model_metric)
single_prs_model_performance = rbind(all_model_performance)
single_prs_model_summary = single_prs_model_metric %>% 
  inner_join(single_prs_model_performance)
```

### 3) Multiple PRS model
```{r}
#========= multiple PRS model =========
# AD PRS combinations
X = c("AD_PRS_EUR_indsig", "AD_PRS_Trans_indsig", "AD_PRS_AFR_indsig", demo_set)
PRS.AD.indsig.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                balance_classes = T, 
                                model_id = "AD PRS ind.sig combine", 
                                nfolds = 5, fold_assignment = "Stratified", 
                                seed = seed_num, family = "binomial", lambda = 0, 
                                standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_lead", "AD_PRS_Trans_lead", "AD_PRS_AFR_lead", demo_set)
PRS.AD.lead.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                              balance_classes = T, 
                              model_id = "AD PRS lead combine", 
                              nfolds = 5, fold_assignment = "Stratified", 
                              seed = seed_num, family = "binomial", lambda = 0, 
                              standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_map", "AD_PRS_Trans_map", "AD_PRS_AFR_map", demo_set)
PRS.AD.map.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                             balance_classes = T, 
                             model_id = "AD PRS map combine", 
                             nfolds = 5, fold_assignment = "Stratified", 
                             seed = seed_num, family = "binomial", lambda = 0, 
                             standardize = T, keep_cross_validation_models = F)
# AD + neuro combinations
X = c("AD_PRS_EUR_indsig", "AD_PRS_Trans_indsig", "AD_PRS_AFR_indsig", 
      "PD_PRS_indsig", "LBD_PRS_indsig", "PSP_PRS_indsig", 
      "Stroke_PRS_indsig", demo_set)
PRS.neuro.indsig.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                   balance_classes = T, 
                                   model_id = "Neuro PRS ind.sig combine", 
                                   nfolds = 5, fold_assignment = "Stratified", 
                                   seed = seed_num, family = "binomial", lambda = 0, 
                                   standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_lead", "AD_PRS_Trans_lead", "AD_PRS_AFR_lead", 
      "PD_PRS_lead", "LBD_PRS_lead", "PSP_PRS_lead", 
      "Stroke_PRS_lead", demo_set)
PRS.neuro.lead.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                 balance_classes = T, 
                                 model_id = "Neuro PRS lead combine", 
                                 nfolds = 5, fold_assignment = "Stratified", 
                                 seed = seed_num, family = "binomial", lambda = 0, 
                                 standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_map", "AD_PRS_Trans_map", "AD_PRS_AFR_map", 
      "PD_PRS_map", "LBD_PRS_map", "PSP_PRS_map", "Stroke_PRS_map", demo_set)
PRS.neuro.map.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                balance_classes = T,
                                model_id = "Neuro PRS map combine", 
                                nfolds = 5, fold_assignment = "Stratified", 
                                seed = seed_num, family = "binomial", lambda = 0, 
                                standardize = T, keep_cross_validation_models = F)
# LASSO selected PRSs
X = c(demo_set, names(target.train)[8:28])
PRS.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = 1, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
prs_select = PRS.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
# Step 2: fit LASSO
X = c(prs_select, demo_set)
PRS.neuro.select = h2o.glm(training_frame = target.train, x = X, y = Y, 
                           balance_classes = T, model_id = "LASSO PRSs", 
                           nfolds = 5, fold_assignment = "Stratified", 
                           seed = seed_num, family = "binomial",
                           alpha = 0, lambda_search = TRUE,
                           standardize = T, keep_cross_validation_models = F)
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("PRS.AD.indsig.combine", "PRS.AD.lead.combine", 
               "PRS.AD.map.combine", "PRS.neuro.indsig.combine", 
               "PRS.neuro.lead.combine", "PRS.neuro.map.combine", 
               "PRS.neuro.select")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
multi_prs_model_metric = rbind(all_model_metric)
multi_prs_model_performance = rbind(all_model_performance)
multi_prs_model_summary = multi_prs_model_metric %>% 
  inner_join(multi_prs_model_performance)

full_prs_model_summary = rbind(single_prs_model_summary, 
                               multi_prs_model_summary) %>% as.data.frame()
```

## 2. Weighted-SNP models
```{r}
snp_df_full = sample_AFR_final %>% 
  mutate(UniqueSampleId = as.character(UniqueSampleId)) %>% 
  left_join(genotype_df_fam, by = c("UniqueSampleId" = "IID")) %>% 
  select(-c(UniqueSampleId)) 
# LASSO selected SNPs
target.train = as.h2o(snp_df_full)
target.train[,Y] = as.factor(target.train[,Y])
# Set parameters
alpha_feature = 0.4
```

### Run two-step linear models
```{r}
# AD indsig SNPs
# Step 1: feature selection
X = c(demo_set, AD.indsig.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AFR_AD_indsig = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AFR_AD_indsig))
# Step 2: fit LASSO
X = c(snps_AFR_AD_indsig, demo_set)
aml_AD_indsig = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.AD.indsig = aml_AD_indsig@leader

# AD lead SNPs
# Step 1: feature selection
X = c(demo_set, AD.lead.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AFR_AD_lead = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AFR_AD_lead))
# Step 2: fit LASSO
X = c(snps_AFR_AD_lead, demo_set)
aml_AD_lead = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.AD.lead = aml_AD_lead@leader

# AD map SNPs
# Step 1: feature selection
X = c(demo_set, AD.map.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AFR_AD_map = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AFR_AD_map))
# Step 2: fit LASSO
X = c(snps_AFR_AD_map, demo_set)
aml_AD_map = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.AD.map = aml_AD_map@leader

# All indsig SNPs
# Step 1: feature selection
X = c(demo_set, all.indsig.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AFR_all_indsig = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AFR_all_indsig))
# Step 2: fit LASSO
X = c(snps_AFR_all_indsig, demo_set)
aml_all_indsig = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.all.indsig = aml_all_indsig@leader

# All lead SNPs
# Step 1: feature selection
X = c(demo_set, all.lead.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AFR_all_lead = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AFR_all_lead))
# Step 2: fit LASSO
X = c(snps_AFR_all_lead, demo_set)
aml_all_lead = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.all.lead = aml_all_lead@leader

# All mapped SNPs
# Step 1: feature selection
X = c(demo_set, all.map.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_AFR_all_map = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_AFR_all_map))
# Step 2: fit LASSO
X = c(snps_AFR_all_map, demo_set)
aml_all_map = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.all.map = aml_all_map@leader
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("snp.lasso.AD.indsig", "snp.lasso.AD.lead", 
               "snp.lasso.AD.map", "snp.lasso.all.indsig",
               "snp.lasso.all.lead", "snp.lasso.all.map")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
lasso_snp_model_metric = rbind(all_model_metric)
lasso_snp_model_performance = rbind(all_model_performance)
lasso_snp_model_summary = lasso_snp_model_metric %>% 
  inner_join(lasso_snp_model_performance)
```

## 3. AutoML
```{r warning=FALSE}
# First use the sample selection results from LASSO 
X = c(snps_AFR_all_map, demo_set)
# Run AutoML for 10 base models (limited to 1 hour max runtime by default)
aml.full.map = h2o.automl(x = X, y = Y, training_frame = target.train, 
                          balance_classes = TRUE, max_models = 10, 
                          seed = seed_num, sort_metric = "AUCPR")
# Optionally edd extra model information to the leaderboard
lb.aml = h2o.get_leaderboard(aml.full.map, extra_columns = "ALL")
# Print all rows (instead of default 6 rows)
print(lb.aml, n = nrow(lb.aml))
```

```{r}
# Get the "best" for each type of model
best_gbm = h2o.getModel("GBM_1_AutoML_253_20230615_231930")
best_xgboost = h2o.getModel("XGBoost_1_AutoML_253_20230615_231930")
best_drf = h2o.getModel("DRF_1_AutoML_253_20230615_231930")
best_xrt = h2o.getModel("XRT_1_AutoML_253_20230615_231930")
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("best_gbm", "best_xgboost", "best_drf", "best_xrt")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
aml_snp_model_metric = rbind(all_model_metric)
aml_snp_model_performance = rbind(all_model_performance)
aml_snp_model_summary = aml_snp_model_metric %>% 
  inner_join(aml_snp_model_performance)
```

```{r}
full_model_summary = rbind(full_prs_model_summary, 
                           lasso_snp_model_summary,
                           aml_snp_model_summary) %>% as.data.frame()
# Output results to table
write.table(full_model_summary, 
            file = paste0(output_path, "Model_performance_logistic_AFR.txt"),
            sep = "\t", quote = F, row.names = F, col.names = T)
# h2o.shutdown(prompt = F)
```


# Part 3. EAS analyses
````{r}
# Read in cleaned data
load(file = paste0(raw_data_path, "modeling/EAS/sample_EAS_final.rda"))
# Join PRSs
sample_EAS_prs = sample_EAS_final %>% 
  mutate(UniqueSampleId = as.character(UniqueSampleId)) %>% 
  left_join(atlas_prs_final) %>% column_to_rownames(var = "UniqueSampleId") %>% 
  select(-c(gen_ancestry, APOE, e2count)) %>% drop_na()
dim(sample_EAS_prs) # dim = (1638,29)
# Set response variable
Y = "dementia"
demo_set = c("age", "female", paste0('PC', 1:4))
# set h2o project
# h2o.init(nthreads = -1)
```

## 1. PRS models
### 1) Model setups
```{r}
target.train = as.h2o(sample_EAS_prs)
target.train[,Y] = as.factor(target.train[,Y])
```

### 2) Single PRS models
```{r}
seed_num = 970103
#========= Benchmark: demo only =========
X = c("age", "female")
demo.only = h2o.glm(training_frame = target.train, x = X, y = Y, 
                    balance_classes = T, model_id = "Demo only", 
                    nfolds = 5, fold_assignment = "Stratified",
                    seed = seed_num, family = "binomial", lambda = 0, 
                    standardize = T, keep_cross_validation_models = F)
#========= single PRS model =========
# AD PRS EUR
X = c("AD_PRS_EUR_indsig", demo_set)
PRS.EUR.indsig = h2o.glm(training_frame = target.train, x = X, y = Y, 
                         balance_classes = T, model_id = "AD PRS EUR indsig", 
                         nfolds = 5, fold_assignment = "Stratified", 
                         seed = seed_num, family = "binomial", lambda = 0, 
                         standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_lead", demo_set)
PRS.EUR.lead = h2o.glm(training_frame = target.train, x = X, y = Y, 
                       balance_classes = T, model_id = "AD PRS EUR lead", 
                       nfolds = 5, fold_assignment = "Stratified", 
                       seed = seed_num, family = "binomial", lambda = 0, 
                       standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_map", demo_set)
PRS.EUR.map = h2o.glm(training_frame = target.train, x = X, y = Y, 
                      balance_classes = T, model_id = "AD PRS EUR map", 
                      nfolds = 5, fold_assignment = "Stratified", 
                      seed = seed_num, family = "binomial", lambda = 0, 
                      standardize = T, keep_cross_validation_models = F)
# AD PRS AFR
X = c("AD_PRS_AFR_indsig", demo_set)
PRS.AFR.indsig = h2o.glm(training_frame = target.train, x = X, y = Y, 
                         balance_classes = T, model_id = "AD PRS AFR indsig", 
                         nfolds = 5, fold_assignment = "Stratified", 
                         seed = seed_num, family = "binomial", lambda = 0, 
                         standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_AFR_lead", demo_set)
PRS.AFR.lead = h2o.glm(training_frame = target.train, x = X, y = Y, 
                       balance_classes = T, model_id = "AD PRS AFR lead", 
                       nfolds = 5, fold_assignment = "Stratified", 
                       seed = seed_num, family = "binomial", lambda = 0, 
                       standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_AFR_map", demo_set)
PRS.AFR.map = h2o.glm(training_frame = target.train, x = X, y = Y, 
                      balance_classes = T, model_id = "AD PRS AFR map", 
                      nfolds = 5, fold_assignment = "Stratified", 
                      seed = seed_num, family = "binomial", lambda = 0, 
                      standardize = T, keep_cross_validation_models = F)
# AD PRS Trans 
X = c("AD_PRS_Trans_indsig", demo_set)
PRS.trans.indsig = h2o.glm(training_frame = target.train, x = X, y = Y, 
                           balance_classes = T, model_id = "AD PRS Trans indsig", 
                           nfolds = 5, fold_assignment = "Stratified", 
                           seed = seed_num, family = "binomial", lambda = 0, 
                           standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_Trans_lead", demo_set)
PRS.trans.lead = h2o.glm(training_frame = target.train, x = X, y = Y, 
                         balance_classes = T, model_id = "AD PRS Trans lead", 
                         nfolds = 5, fold_assignment = "Stratified", 
                         seed = seed_num, family = "binomial", lambda = 0, 
                         standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_Trans_map", demo_set)
PRS.trans.map = h2o.glm(training_frame = target.train, x = X, y = Y, 
                        balance_classes = T, model_id = "AD PRS Trans map", 
                        nfolds = 5, fold_assignment = "Stratified", 
                        seed = seed_num, family = "binomial", lambda = 0, 
                        standardize = T, keep_cross_validation_models = F)
# e4count
X = c("e4count", demo_set)
e4count = h2o.glm(training_frame = target.train, x = X, y = Y, 
                  balance_classes = T, model_id = "e4count", 
                  nfolds = 5, fold_assignment = "Stratified", 
                  seed = seed_num, family = "binomial", lambda = 0, 
                  standardize = T, keep_cross_validation_models = F)
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("demo.only", "e4count", 
               "PRS.EUR.indsig", "PRS.EUR.lead", "PRS.EUR.map", 
               "PRS.AFR.indsig", "PRS.AFR.lead", "PRS.AFR.map", 
               "PRS.trans.indsig", "PRS.trans.lead", "PRS.trans.map")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>%
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
single_prs_model_metric = rbind(all_model_metric)
single_prs_model_performance = rbind(all_model_performance)
single_prs_model_summary = single_prs_model_metric %>% 
  inner_join(single_prs_model_performance)
```

### 3) Multiple PRS model
```{r}
#========= multiple PRS model =========
# AD PRS combinations
X = c("AD_PRS_EUR_indsig", "AD_PRS_Trans_indsig", "AD_PRS_AFR_indsig", demo_set)
PRS.AD.indsig.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                balance_classes = T, 
                                model_id = "AD PRS ind.sig combine", 
                                nfolds = 5, fold_assignment = "Stratified", 
                                seed = seed_num, family = "binomial", lambda = 0, 
                                standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_lead", "AD_PRS_Trans_lead", "AD_PRS_AFR_lead", demo_set)
PRS.AD.lead.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                              balance_classes = T, 
                              model_id = "AD PRS lead combine", 
                              nfolds = 5, fold_assignment = "Stratified", 
                              seed = seed_num, family = "binomial", lambda = 0, 
                              standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_map", "AD_PRS_Trans_map", "AD_PRS_AFR_map", demo_set)
PRS.AD.map.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                             balance_classes = T, 
                             model_id = "AD PRS map combine", 
                             nfolds = 5, fold_assignment = "Stratified", 
                             seed = seed_num, family = "binomial", lambda = 0, 
                             standardize = T, keep_cross_validation_models = F)
# AD + neuro combinations
X = c("AD_PRS_EUR_indsig", "AD_PRS_Trans_indsig", "AD_PRS_AFR_indsig", 
      "PD_PRS_indsig", "LBD_PRS_indsig", "PSP_PRS_indsig", 
      "Stroke_PRS_indsig", demo_set)
PRS.neuro.indsig.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                   balance_classes = T, 
                                   model_id = "Neuro PRS ind.sig combine", 
                                   nfolds = 5, fold_assignment = "Stratified", 
                                   seed = seed_num, family = "binomial", lambda = 0, 
                                   standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_lead", "AD_PRS_Trans_lead", "AD_PRS_AFR_lead", 
      "PD_PRS_lead", "LBD_PRS_lead", "PSP_PRS_lead", 
      "Stroke_PRS_lead", demo_set)
PRS.neuro.lead.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                 balance_classes = T, 
                                 model_id = "Neuro PRS lead combine", 
                                 nfolds = 5, fold_assignment = "Stratified", 
                                 seed = seed_num, family = "binomial", lambda = 0, 
                                 standardize = T, keep_cross_validation_models = F)
X = c("AD_PRS_EUR_map", "AD_PRS_Trans_map", "AD_PRS_AFR_map", 
      "PD_PRS_map", "LBD_PRS_map", "PSP_PRS_map", "Stroke_PRS_map", demo_set)
PRS.neuro.map.combine = h2o.glm(training_frame = target.train, x = X, y = Y, 
                                balance_classes = T,
                                model_id = "Neuro PRS map combine", 
                                nfolds = 5, fold_assignment = "Stratified", 
                                seed = seed_num, family = "binomial", lambda = 0, 
                                standardize = T, keep_cross_validation_models = F)
# LASSO selected PRSs
X = c(demo_set, names(target.train)[8:28])
PRS.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = 1, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
prs_select = PRS.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
# Step 2: fit LASSO
X = c(prs_select, demo_set)
PRS.neuro.select = h2o.glm(training_frame = target.train, x = X, y = Y, 
                           balance_classes = T, model_id = "LASSO PRSs", 
                           nfolds = 5, fold_assignment = "Stratified", 
                           seed = seed_num, family = "binomial",
                           alpha = 0, lambda_search = TRUE,
                           standardize = T, keep_cross_validation_models = F)
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("PRS.AD.indsig.combine", "PRS.AD.lead.combine", 
               "PRS.AD.map.combine", "PRS.neuro.indsig.combine", 
               "PRS.neuro.lead.combine", "PRS.neuro.map.combine", 
               "PRS.neuro.select")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
multi_prs_model_metric = rbind(all_model_metric)
multi_prs_model_performance = rbind(all_model_performance)
multi_prs_model_summary = multi_prs_model_metric %>% 
  inner_join(multi_prs_model_performance)

full_prs_model_summary = rbind(single_prs_model_summary, 
                               multi_prs_model_summary) %>% as.data.frame()
```

## 2. Weighted-SNP models
```{r}
snp_df_full = sample_EAS_final %>% 
  mutate(UniqueSampleId = as.character(UniqueSampleId)) %>% 
  left_join(genotype_df_fam, by = c("UniqueSampleId" = "IID")) %>% 
  select(-c(UniqueSampleId)) 
# LASSO selected SNPs
target.train = as.h2o(snp_df_full)
target.train[,Y] = as.factor(target.train[,Y])
# Set parameters
alpha_feature = 0.6
```

### Run two-step linear models
```{r}
# AD indsig SNPs
# Step 1: feature selection
X = c(demo_set, AD.indsig.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_EAS_AD_indsig = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_EAS_AD_indsig))
# Step 2: fit LASSO
X = c(snps_EAS_AD_indsig, demo_set)
aml_AD_indsig = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.AD.indsig = aml_AD_indsig@leader

# AD lead SNPs
# Step 1: feature selection
X = c(demo_set, AD.lead.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_EAS_AD_lead = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_EAS_AD_lead))
# Step 2: fit LASSO
X = c(snps_EAS_AD_lead, demo_set)
aml_AD_lead = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.AD.lead = aml_AD_lead@leader

# AD map SNPs
# Step 1: feature selection
X = c(demo_set, AD.map.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_EAS_AD_map = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_EAS_AD_map))
# Step 2: fit LASSO
X = c(snps_EAS_AD_map, demo_set)
aml_AD_map = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.AD.map = aml_AD_map@leader

# All indsig SNPs
# Step 1: feature selection
X = c(demo_set, all.indsig.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_EAS_all_indsig = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_EAS_all_indsig))
# Step 2: fit LASSO
X = c(snps_EAS_all_indsig, demo_set)
aml_all_indsig = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.all.indsig = aml_all_indsig@leader

# All lead SNPs
# Step 1: feature selection
X = c(demo_set, all.lead.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_EAS_all_lead = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_EAS_all_lead))
# Step 2: fit LASSO
X = c(snps_EAS_all_lead, demo_set)
aml_all_lead = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.all.lead = aml_all_lead@leader

# All mapped SNPs
# Step 1: feature selection
X = c(demo_set, all.map.combine.id)
SNP.lasso.feature = h2o.glm(training_frame = target.train, x = X, y = Y, 
                            balance_classes = T, model_id = "LASSO features", 
                            nfolds = 5, fold_assignment = "Stratified", 
                            seed = seed_num, family = "binomial", 
                            alpha = alpha_feature, lambda_search = TRUE, standardize = T, 
                            keep_cross_validation_models = F)
snps_EAS_all_map = SNP.lasso.feature@model$variable_importances %>% 
  as.data.frame() %>% filter(variable %!in% demo_set) %>% 
  filter(scaled_importance > 0) %>% pull(variable)
print(length(snps_EAS_all_map))
# Step 2: fit LASSO
X = c(snps_EAS_all_map, demo_set)
aml_all_map = h2o.automl(x = X, y = Y, training_frame = target.train, 
                           balance_classes = TRUE, max_models = 10, nfolds = 5, 
                           seed = seed_num, sort_metric = "AUCPR", 
                           include_algos = "GLM")
snp.lasso.all.map = aml_all_map@leader
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("snp.lasso.AD.indsig", "snp.lasso.AD.lead", 
               "snp.lasso.AD.map", "snp.lasso.all.indsig",
               "snp.lasso.all.lead", "snp.lasso.all.map")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
lasso_snp_model_metric = rbind(all_model_metric)
lasso_snp_model_performance = rbind(all_model_performance)
lasso_snp_model_summary = lasso_snp_model_metric %>% 
  inner_join(lasso_snp_model_performance)
```

## 3. AutoML
```{r warning=FALSE}
# First use the sample selection results from LASSO 
X = c(snps_EAS_all_map, demo_set)
# Run AutoML for 10 base models (limited to 1 hour max runtime by default)
aml.full.map = h2o.automl(x = X, y = Y, training_frame = target.train, 
                          balance_classes = TRUE, max_models = 10, 
                          seed = seed_num, sort_metric = "AUCPR")
# Optionally edd extra model information to the leaderboard
lb.aml = h2o.get_leaderboard(aml.full.map, extra_columns = "ALL")
# Print all rows (instead of default 6 rows)
print(lb.aml, n = nrow(lb.aml))
```

```{r}
# Get the "best" for each type of model
best_gbm = h2o.getModel("GBM_1_AutoML_260_20230615_233155")
best_xgboost = h2o.getModel("XGBoost_2_AutoML_260_20230615_233155")
best_drf = h2o.getModel("DRF_1_AutoML_260_20230615_233155")
best_xrt = h2o.getModel("XRT_1_AutoML_260_20230615_233155")
```

```{r}
all_model_metric = c()
all_model_performance = c()
# Extract results
model_list = c("best_gbm", "best_xgboost", "best_drf", "best_xrt")
for (i in 1:length(model_list)) {
  print(paste0("Model: ", model_list[i]))
  # Run extraction function
  extract_cv = extract_cv_results(get(model_list[i]))
  model_metric = t(extract_cv[[1]]) %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  model_performance = extract_cv[[2]] %>% as.data.frame() %>% 
    mutate(mark = model_list[i])
  if (i == 1) {
    all_model_metric = model_metric
    all_model_performance = model_performance
  } else {
    all_model_metric = rbind(all_model_metric, model_metric)
    all_model_performance = rbind(all_model_performance, model_performance)
  }
}
# Join results together
aml_snp_model_metric = rbind(all_model_metric)
aml_snp_model_performance = rbind(all_model_performance)
aml_snp_model_summary = aml_snp_model_metric %>% 
  inner_join(aml_snp_model_performance)
```

```{r}
full_model_summary = rbind(full_prs_model_summary, 
                           lasso_snp_model_summary,
                           aml_snp_model_summary) %>% as.data.frame()
# Output results to table
write.table(full_model_summary, 
            file = paste0(output_path, "Model_performance_logistic_EAS.txt"),
            sep = "\t", quote = F, row.names = F, col.names = T)
# h2o.shutdown(prompt = F)
```

# Part 5. LASSO interation
```{r}
random_seed_lst = sample(1:1000000, 1000, replace = FALSE)
Y = "dementia"
# LASSO selected SNPs
target.train = as.h2o(snp_df_full)
target.train[,Y] = as.factor(target.train[,Y])
X = c(demo_set, all.map.combine.id)
for(i in 1:1000) {
  seed_num = random_seed_lst[i]
  print(paste0("Iteration: ", i))
  set.seed(seed_num)
  SNP.glm.lasso = h2o.glm(training_frame = target.train, x = X, y = Y, 
                        balance_classes = T, model_id = "LASSO SNPs", 
                        nfolds = 5, fold_assignment = "Stratified", 
                        seed = seed_num, family = "binomial", 
                        alpha = 1, lambda_search = TRUE, standardize = T, 
                        keep_cross_validation_models = F)
  var_imp = SNP.glm.lasso@model$variable_importances %>% as.data.frame() %>% 
    filter(scaled_importance > 0) %>% select(variable, percentage)
  if (i == 1) {
    lasso_vars_snps = var_imp
  } else {
    lasso_vars_snps = rbind(lasso_vars_snps, var_imp)
  }
}
selected_snp_summary_allmap = lasso_vars_snps %>% group_by(variable) %>% 
  mutate(freq = n(), avg_percentage = sum(percentage, na.rm = T)*100/1000) %>% 
  select(variable, freq, avg_percentage) %>% unique() %>% 
  arrange(desc(freq))
save(selected_snp_summary_allmap, 
     file = paste0(raw_data_path, "modeling/AMR/lasso_snp_allmap_0614.rda"))
```